---
title: "Energy Distribution"
author: "Sara Orofino"
date: "10/3/2019"
output: html_document
---

We can use this rmd to determine the latitudinal distribution of each energy mix. I clarifed with Lee and he said each mix of energy should be assigned across the globe 1,000 different ways with weights assigned based on suitability of the area for providing that type of energy. 


First we need to get the rasters from everyone for global potential
read in using readTIFF(source, native = TRUE) - native will turn it into a native raster we can plot

```{r packages}
library(sp)
#library(tiff)
library(raster)
#library(rgdal)
```

```{r read_data}
# solar_raster <- system.file("~Documents/Github/Climate-Change-Biology/solar_potential.tif", package="raster")
#This doesnt work... it shows up in the environment but its empty...

# JKM edits
# changed system.file to raster
# removed file pathway
# removed package = "raster"
# moved tif file to folder "Energy_Distribution_Calculations" because none of the pathway designations were working. 
solar_raster <- raster("solar_potential.tif")
solar_raster
plot(solar_raster)
```

(Nice job Jamie!)

Here's exactly Patrick's code for how to do this just using this solar raster instead of his example raster:
```{r solar_example}

#make a raster defining the desired inclusion probabilities 
#for the all locations available for sampling
probrast<-solar_raster

#inclusion probability for cells with value >=400 
#will be 10 times that for cells with value <400
## would probably skip this step unless there are real reasons to make the probabilities categorical like this
## DEFINITELY make sure areas with no potential are 0 
probrast[solar_raster>=400]<-10 
probrast[solar_raster<400]<-1

#normalise the probability raster by dividing by the sum of all inclusion weights:
probrast<-probrast/sum(getValues(probrast), na.rm=T)

#plot the raster of inclusion probabilities
plot(probrast)

####### Jamie I don't think this worked the same - it looks like all the values are zero; might have to do with the 400 but I'm not sure where that value came from.

### get all the values in your probability raster as a vector 
 x<-getValues(probrast)
 #set NA cells in raster to zero
 x[is.na(x)]<-0
 
 ### sample from a list of cell IDs from 1:totala cells in raster -- size is number in sample, and prob is the probability of selecting any cell (values from probrast)
 samp<-sample(nrow(probrast)*ncol(probrast), size=100, prob=x)
 
## make a new raster to see what you sampled 
 samprast<-probrast
 
 ## brakets say turn all the cell IDs identified by 'samp' into 1
 samprast[samp]<-1 
 
 ## see how the sampled raster compares to the prob raster
 plot(samprast)
 s <- stack(probrast,samprast)
 plot(s)

 ### this will check to see how many cells from your sample you need to make your energy target
 ## set threshold for TWh needed - 113,009 TWh
 ## Based on  "Key World Energy Statistics 2019". IEA. 26 September 2019. Retrieved 27 October 2019.
 thresh <- 113009
 
 ### get the values from your energy potential map
 v <- getValues(solar_raster)
 
 ## subset those values according to your sample
 v <- v[samp]
 
 ## get a running total of summed energy delivered
 v.sum <- cumsum(v)
 
 ## check to see where in your sample you achieve the target
 stop <- min(which(v.sum>thresh))
  
 ### make a new raster for final selected cells
 selectrast<-solar_raster
 
 ### set all cells in your sample (up to where you achieve energy target to 1)
 selectrast[samp[1:stop]]<-1 
  
## see how it all looks!
plot(selectrast)
s2 <- stack(probrast,samprast,selectrast)
plot(s2)


# So the code runs but I don't know if this makes any sense...
```

