---
title: "Energy Distribution"
author: "Sara Orofino"
date: "10/3/2019"
output: html_document
---

We can use this rmd to determine the latitudinal distribution of each energy mix. I clarifed with Lee and he said each mix of energy should be assigned across the globe 1,000 different ways with weights assigned based on suitability of the area for providing that type of energy. 


First we need to get the rasters from everyone for global potential
read in using readTIFF(source, native = TRUE) - native will turn it into a native raster we can plot

```{r packages}
library(sp)
#library(tiff)
library(raster)
#library(rgdal)
```

```{r read_data}
# solar_raster <- system.file("~Documents/Github/Climate-Change-Biology/solar_potential.tif", package="raster")
#This doesnt work... it shows up in the environment but its empty...

# JKM edits
# changed system.file to raster
# removed file pathway
# removed package = "raster"
# moved tif file to folder "Energy_Distribution_Calculations" because none of the pathway designations were working. 
solar_raster <- raster("solar_potential.tif")
solar_raster
plot(solar_raster)
```


```{r read_data}
#wind raster, 100m hieght

wind_raster <- raster("gwa3_250_power-density_100m.tif")
wind_raster
plot(wind_raster)
```
