---
title: "Energy Mix"
author: "Sara Orofino"
date: "10/3/2019"
output: html_document
---
Packages:
```{r packages}
library(tidyverse)
```

I think it's probably a good idea to try and keep things separate so the work flow is easier to follow. We can use this rmd to figure out a code for how to select an energy mix that meets global demand. 

There are seven kinds of energy to choose from:
 - Carbon Capture and Storage (cc)
 - Solar (s)
 - Wind (w)
 - Solar and wind co-location (sw)
 - Geothermal (g)
 - Biofuel (b)
 - Nuclear (n)
 
Goal: to meet global demand (demand) by randomly selecting a mix of energy from the seven sources above using weights for how much energy is likely to be contributed by each source. 

Parameters:
demand - Global energy demand (TWh); determined from literature (let's say 23000 TWh for now)


Total Energy Demand equation: The total energy demand (for example 23,000 TWh or 100%) is the sum of the amount (or percent) of energy from the seven different types
```{r}

# Function for total energy demand
demand <- function(cc, s, w, sw, g, b, n) {
  demand <- cc + s + + w + sw + g + b + n
}

```

How do we determine how much energy comes from each source?

Maybe we can use random number generators with a mean and standard deviation from the literature for how much energy could be supplied from each source. Alternatively, use a random number generator with a mean of 0 and standard deviation of 1 and multiply by the weight for each energy type.
 
For example let's use solar
Say that solar and wind energy has the potential to contribute (on average) 7% of global energy demand or 1610 TWh and has a standard deviation of 257 TWh
 
```{r}
# To calculate a numeric value - use the mean and standard deviation (this is made up data)
sw_1 <- rnorm(1, 1610, 257)

#To calculate a percent - use 7% and the standard deviation (this is made up data)
sw_2 <- rnorm(1, 7, 1)
```
 
Is there a way to generate seven numbers that all add up to 100? (AKA different %s of energy that add up to 100% of demand)

**Okay let's try something different**

If we know what percent of global demand each source could potentially match we can determine how many TWh should be in each energy "bucket". We then draw 23000 TWh at random from the seven different buckets. 

For example here's some made up numbers on the percent of global demand each energy source could meet:
- Carbon Capture and Storage (30%)
 - Solar (50%)
 - Wind (50%)
 - Solar and wind co-location (10%)
 - Geothermal (15%)
 - Biofuel (20%)
 - Nuclear (20%)
 
 We could first determine how much TWh of energy those percentages equate to - 
```{r}
cc <- 23000 * 0.30
s <- 23000 * 0.50
w <- 23000 * 0.50
sw <- 23000 * 0.10
g <- 23000 * 0.15
b <- 23000 * 0.20
n <- 23000 * 0.20

# Acutally this part might be unnecessary
```
 
Then we need to draw 23000 TWh of energy (one TWh at a time) based on the probabilities above. I think we might be able to do this using sample()

```{r}

# First create a vector of elements to sample from (in this case all of our energy types)
energy <- c("cc", "s", "w", "sw", "g", "b", "n")

#Assign n the value of how many samples we want to take (we need to draw 23000 energy types if each one represents just 1 TWh)
n <- 23000 

#Create a vector of the probabilities associated with drawing each type of energy (this comes from the energy groups)
## Important to enter the probabilities in the same order as the energy types are listed 
probs <- c(0.30, 0.50, 0.50, 0.10, 0.15, 0.20, 0.20)

#Now run the sample function:
trial_1 <- sample(energy, size = n, replace = TRUE, prob = probs)

#Convert to a dataframe:
trial_1 <- as.data.frame(trial_1)

#See if these results make sense - how many times was each type of energy chosen?
sum(trial_1$trial_1 == "cc") #3579
sum(trial_1$trial_1 == "s") #5921
sum(trial_1$trial_1 == "w") #5857
sum(trial_1$trial_1 == "sw") #1205
sum(trial_1$trial_1 == "g") #1768
sum(trial_1$trial_1 == "b") #2354
sum(trial_1$trial_1 == "n") #2316

#Solar and wind were selected the most and had the highest probabilities; sw was the least and had the lowest probabilities. Solar and wind had the same probabilities and the numbers were similar; biofuel and nuclear had the same probabilities and the numbers were also similar. This method seems to make sense. 
```

Now is there a way to tally these that is faster than summing each one? Not sure if we need to do this but if its short and quick could be useful to know. 
```{r}

#Group by energy type and count how many in each category - mutate to add the percent of the total 
trial1_tally <- trial_1 %>% 
  group_by(trial_1) %>% 
  tally() %>% 
  rename(energy_type = "trial_1", TWh = "n") %>% 
  mutate(percent_of_demand = round(((TWh/23000)*100), 2))

```

Cool- now loop it....
```{r}
#Write the sample() code as a function but leave it in long form where every energy piece is its own row
## Aka 23000 observations of 1 variable 
energy_mixes_long <- function(energy){
  out <- sample(energy, size = n, replace = TRUE, prob = probs)
  return(as.data.frame(out) %>% 
           rename(energy_type = "out"))
}

#Write the sample() code as a function and add the wranggling to condense the dataframe down
## Aka 7 observations (1 for each energy type), total TWh for selected for that energy type, and % of the total energy demand
energy_mixes_short <- function(energy){
  out <- sample(energy, size = n, replace = TRUE, prob = probs)
  return(as.data.frame(out) %>% 
           rename(energy_type = "out") %>%
           group_by(energy_type) %>% 
           tally() %>% 
           rename(TWh = "n") %>% 
           mutate(percent_of_demand = round(((TWh/23000)*100), 2)))
}

#Example:
mix1_long <- energy_mixes_long(energy)
mix1_short <- energy_mixes_short(energy)

mix2_long <- energy_mixes_long(energy)
mix2_short <- energy_mixes_short(energy)

# This is cool this seems to be working - but I don't know how to loop it so that every data frame has its own name

```

Okay trying the things Patrick suggested:

Method 1 - Use a range of the expected values for each source. For example if solar has an expected energy production between 5-25% then we use sample(5:25,1).

```{r}

###### Using only 10 draws as an example to see if this works:

# set up a results data frame to write successful results to (using 10 for trial purposes)
results <- data.frame(matrix(ncol=7,nrow=10))

## set a counter variable
x <-1 

## set up a repeater function so it goes until a condition is met
repeat({

## sample integer between 0 and 100 -- if certain energy sources have limits can define the range in the sample() function
## e.g. if solar has to be between 10 and 30% can do solar<-sample(10:30,1)
cc <- sample(0:10,1)
s <- sample(0:50,1)
w <- sample(0:20,1)
sw <- sample(0:30,1)
g <- sample(0:30,1)
b <- sample(0:40,1)
n <- sample(0:40,1)

## find the sum of the mix
s <- c(cc, s, w, sw, g, b, n)
ssum <- sum(s)

## write the mix if it equals 100
if (ssum==100){
results[x,1:7]<-s

## increase the counter to keep track of successful mixes
x<-x+1}

## break after 100 mixes written
if (x>10){
break}

})

## write the mixes to a file 
write.csv(results, "trial_energymix.csv",row.names=FALSE)
```


Method 2 Centered toward the expected range. For example, say again that solar has an expected energy production range between 5-25%. We want to weight the samples so they fall closer to the mean of this expected range. 

```{r}
##### Using only 10 draws as an example to see if this works:

## set up a results data frame to write successful results to (using 10 for trial purposes)
results2 <- data.frame(matrix(ncol=7,nrow=10))

## set a counter variable
x <-1 

## set up a repeater function so it goes until a condition is met
repeat({

# Set range values for each type of energy:
#### Carbon Capture 0-10
minVal_cc <- 0
maxVal_cc <- 10
count <- 1
mn_cc <- (maxVal_cc + minVal_cc)/2
sd_cc <- (maxVal_cc-minVal_cc)/3  ### increase or decrease this denominator to adjust the shape of the distribution -- larger denominator will be narrower distribution (more values closer to mean)

cc <- rnorm(count, mean = mn_cc, sd = sd_cc)

#get rid of values outside the range if you want -- and round to integer
cc[cc<minVal_s]<- minVal_cc
cc[cc>maxVal_s]<- maxVal_cc
cc <- round(cc, digits=0)

#### Solar 5-25
minVal_s <- 5
maxVal_s <- 25
count <- 1
mn_s <- (maxVal_s + minVal_s)/2
sd_s <- (maxVal_s-minVal_s)/3  ### increase or decrease this denominator to adjust the shape of the distribution -- larger denominator will be narrower distribution (more values closer to mean)

s <- rnorm(count, mean = mn_s, sd = sd_s)

#get rid of values outside the range if you want -- and round to integer
s[s<minVal_s]<- minVal_s
s[s>maxVal_s]<- maxVal_s
s <- round(s, digits=0)

#### Wind 20-50
minVal_w <- 20
maxVal_w <- 50
count <- 1
mn_w <- (maxVal_w + minVal_w)/2
sd_w <- (maxVal_w-minVal_w)/3  ### increase or decrease this denominator to adjust the shape of the distribution -- larger denominator will be narrower distribution (more values closer to mean)

w <- rnorm(count, mean = mn_w, sd = sd_w)

#get rid of values outside the range if you want -- and round to integer
w[w<minVal_w]<- minVal_w
w[w>maxVal_w]<- maxVal_w
w <- round(w, digits=0)

#### Solar/Wind colocation 5-10
minVal_sw <- 5
maxVal_sw <- 10
count <- 1
mn_sw <- (maxVal_sw + minVal_sw)/2
sd_sw <- (maxVal_sw-minVal_sw)/3  ### increase or decrease this denominator to adjust the shape of the distribution -- larger denominator will be narrower distribution (more values closer to mean)

sw <- rnorm(count, mean = mn_sw, sd = sd_sw)

#get rid of values outside the range if you want -- and round to integer
sw[sw<minVal_sw]<- minVal_sw
sw[sw>maxVal_sw]<- maxVal_sw
sw <- round(sw, digits=0)

#### Geothermal 15-35
minVal_g <- 15
maxVal_g <- 35
count <- 1
mn_g <- (maxVal_g + minVal_g)/2
sd_g <- (maxVal_g-minVal_g)/3  ### increase or decrease this denominator to adjust the shape of the distribution -- larger denominator will be narrower distribution (more values closer to mean)

g <- rnorm(count, mean = mn_g, sd = sd_g)

#get rid of values outside the range if you want -- and round to integer
g[g<minVal_g]<- minVal_g
g[g>maxVal_g]<- maxVal_g
g <- round(g, digits=0)

#### Biofuel 20-50
minVal_b <- 20
maxVal_b <- 50
count <- 1
mn_b <- (maxVal_b + minVal_b)/2
sd_b <- (maxVal_b-minVal_b)/3  ### increase or decrease this denominator to adjust the shape of the distribution -- larger denominator will be narrower distribution (more values closer to mean)

b <- rnorm(count, mean = mn_b, sd = sd_b)

#get rid of values outside the range if you want -- and round to integer
b[b<minVal_b]<- minVal_b
b[b>maxVal_b]<- maxVal_b
b <- round(b, digits=0)

#### Nuclear 20-40
minVal_n <- 20
maxVal_n <- 40
count <- 1
mn_n <- (maxVal_n + minVal_n)/2
sd_n <- (maxVal_n-minVal_n)/3  ### increase or decrease this denominator to adjust the shape of the distribution -- larger denominator will be narrower distribution (more values closer to mean)

n <- rnorm(count, mean = mn_n, sd = sd_n)

#get rid of values outside the range if you want -- and round to integer
n[n<minVal_n]<- minVal_n
n[n>maxVal_n]<- maxVal_n
n <- round(n, digits=0)

## find the sum of the mix
sm <- c(cc, s, w, sw, g, b, n)
ssum <- sum(sm)

## write the mix if it equals 100
if (ssum==100){
results2[x,1:7]<-sm

## increase the counter to keep track of successful mixes
x<-x+1}

## break after 10 mixes written
if (x>10){
break}

})
```






