---
title: "Energy Mix"
author: "Sara Orofino"
date: "10/3/2019"
output: html_document
---

I think it's probably a good idea to try and keep things separate so the work flow is easier to follow. We can use this rmd to figure out a code for how to select energy mixes from a (normal?) distribution weighted by how much energy they are likely to provide. 

There are seven kinds of energy to choose from:
 - Carbon Capture and Storage (cc)
 - Solar (s)
 - Wind (w)
 - Solar and wind co-location (sw)
 - Geothermal (g)
 - Biofuel (b)
 - Nuclear (n)
 
Goal: to meet global demand (demand) by randomly selecting a mix of energy from the seven sources above using weights for how much energy is likely to be contributed by each source. 

Parameters:
demand - Global energy demand (TWh); determined from literature (more than 23,000 TWh?)
weights for each of the seven energy types - determined by the different groups researching each energy type (wt_energytype)


Total Energy Demand equation: The total energy demand (for example 23,000 TWh or 100%) is the sum of the amount (or percent) of energy from the seven different types
```{r}

# Function for total energy demand
demand <- function(cc, s, w, sw, g, b, n) {
  demand <- cc + s + + w + sw + g + b + n
}

```

How do we determine how much energy comes from each source?

Maybe we can use random number generators with a mean and standard deviation from the literature for how much energy could be supplied from each source. Alternatively, use a random number generator with a mean of 0 and standard deviation of 1 and multiply by the weight for each energy type.
 
For example let's use solar
Say that solar and wind energy has the potential to contribute (on average) 7% of global energy demand or 1610 TWh and has a standard deviation of 257 TWh
 
```{r}
# To calculate a numeric value - use the mean and standard deviation (this is made up data)
sw_1 <- rnorm(1, 1610, 257)

#To calculate a percent - use 7% and the standard deviation (this is made up data)
sw_2 <- rnorm(1, 7, 1)
```
 
Is there a way to generate seven numbers that all add up to 100? (AKA different %s of energy that add up to 100% of demand)








