---
title: "Model_run"
author: "Sara Orofino"
date: "10/29/2019"
output: html_document
---

```{r packages, include = FALSE}
#packages
library(tidyverse)
library(raster)

```

Alright here we go with the running of this model...

Step 1: Energy mixes - DO NOT RUN
```{r mix_1}
# set up a results data frame to write successful results 
results <- data.frame(matrix(ncol=4,nrow=1000))

## set a counter variable
x <-1 

## set up a repeater function so it goes until a condition is met
repeat({
  
  ## scale up current percentage of energy by sector in 2017 to be able to meet 100% of energy demand - refer to table in methods for the 2017 percentages 
  s <- sample(0:3,1)
  w <- sample(0:7,1)
  b <- sample(0:70,1)
  h <- sample(0:26,1)
  
  ## find the sum of the mix
  s <- c(s, w, b, h)
  ssum <- sum(s)
  
  ## write the mix if it equals 100
  if (ssum==100){
    results[x,1:4]<-s
    
    ## increase the counter to keep track of successful mixes
    x<-x+1}
  
  ## break after 100 mixes written
  if (x>1000){
    break}
  
})

#Now mutate the dataframe to display TWh of energy instead of percents
mix_1 <- results %>% 
  mutate(solar = (X1/ 100) * 153596,
         wind = (X2/ 100) * 153596,
         hydro = (X3/ 100) * 153596,
         biofuel = (X4/ 100) * 153596) %>% 
  dplyr::select(5:8)

#Write csv for later:
#write.csv(mix1, "mix1.csv", row.names = FALSE)
```
These results seem pretty similar let's try another one where we vary the energy sources more even if that's unrealistic. 

```{r mix_2}
### Second set of 1000 mixes 
# set up a results data frame to write successful results 
results_2 <- data.frame(matrix(ncol=4,nrow=1000))

## set a counter variable
x <-1 

## set up a repeater function so it goes until a condition is met
repeat({
  
  ## increase the potential of all energy types to try and diversify the mixes. Make every type 5x higher (up to 100 maximum)
  s <- sample(0:15,1)
  w <- sample(0:35,1)
  b <- sample(0:100,1)
  h <- sample(0:100,1)
  
  ## find the sum of the mix
  s <- c(s, w, b, h)
  ssum <- sum(s)
  
  ## write the mix if it equals 100
  if (ssum==100){
    results_2[x,1:4]<-s
    
    ## increase the counter to keep track of successful mixes
    x<-x+1}
  
  ## break after 100 mixes written
  if (x>1000){
    break}
  
})

#Now mutate the dataframe to display TWh of energy instead of percents
mix_2 <- results_2 %>% 
  mutate(solar = (X1/ 100) * 153596,
         wind = (X2/ 100) * 153596,
         hydro = (X3/ 100) * 153596,
         biofuel = (X4/ 100) * 153596) %>% 
  dplyr::select(5:8)

#Write csv
write.csv(mix_2, "mix2.csv", row.names = FALSE)
```

The second mix has way more variation. Might be good to use both if we get the chance.

Final result of Step 1 is two energy mix dataframes each containing 1000 mixes, the number displayed is the number of TWh contributed by each energy source. This is the energy that needs to be "placed" on the maps in Step 3. 

Step 2: Probability raster maps
```{r read_in}
# Read in rasters:
solar_raster <- raster("solar_potential.tif")
#wind_raster <- raster("wind_raster.tif")
hydro_raster <- raster("hydro_raster.tif")

#mix1 <- read_csv("mix1.csv")
mix2 <- read_csv("mix2.csv") 
hydro=mix2[,3]
max_hydro=cellStats(hydro_raster,sum)/10^9
hydro[hydro>max_hydro]=max_hydro-1000
mix2[,3]=hydro

# Check extents of rasters for masking:
extent(solar_raster) #xmin-max -180 - 180; ymin-max -55 - 60
extent(hydro_raster) #xmin-max -180 - 180; ymin-max -90 - 90

#Change extent of hydro to match solar (the res is also different so use resample)
#hydro_raster <- resample(hydro_raster, solar_raster, resample = 'bilinear')
solar_raster <- disaggregate(solar_raster, fact=2)
solar_raster=solar_raster/4

ext_s <- extent(solar_raster)
hydro_raster <- crop(hydro_raster,ext_s)

```


Only need to create the probability raster map once for each energy source.
```{r prob_rasters}

############### Solar

## make a raster defining the desired inclusion probabilities 
## for the all locations available for sampling
probrast_s<-solar_raster

## normalise the probability raster by dividing by the sum of all inclusion weights:
probrast_s<-probrast_s/sum(getValues(probrast_s), na.rm=T)
probrast_s=probrast_s^6

## plot the raster of inclusion probabilities
plot(probrast_s)

############## Wind

#probrast_w<-wind_raster

#probrast_w<-probrast_w/sum(getValues(probrast_w), na.rm=T)

#plot(probrast_w)

############## Hydro

probrast_h<-hydro_raster

probrast_h<-probrast_h/sum(getValues(probrast_h), na.rm=T)

plot(probrast_h)

############# Biofuel

#probrast_b<-biofuel_raster

#probrast_b<-probrast_b/sum(getValues(probrast_b), na.rm=T)

#plot(probrast_b)
```


Step 3: Distribute the Energy for MIX 2 (because its more interesting and time is limited)
```{r energy_dist}

### Beginning a loop to distribute energy around the globe:

for(i in 1:5){
#{ i=1 
  #Set thresholds for each type based on the TWh from the mixes
  thresh.solar <- mix2[i,1]
  thresh.wind <- mix2[i,2]
  thresh.hydro <- mix2[i,3]
  thresh.bio <- mix2[i,4]
  
  #thresh.solar=thresh.solar*10^9
  #thresh.hydro=thresh.hydro*10^9
  
  ##### Start with solar (s):
  
 xs<-getValues(probrast_s) #get all the values in your probability raster as a vector
 
 xs[is.na(xs)]<-0 #set NA cells in raster to zero
 
 ncells <- length(which(xs!=0))
 ncells
 ### sample from a list of cell IDs from 1:total cells in raster -- size is number in sample, and prob is the probability of selecting any cell (values from probrast)
 samp_s<-sample(nrow(probrast_s)*ncol(probrast_s), size=ncells, prob=xs)
 length(samp_s)
 
 #samprast_s<-probrast_s #make a new raster to see what was sampled
 
 #samprast_s[samp_s]<-1 #brackets say turn all the cell IDs identified by 'samp' into 1

 thresh_s <- thresh.solar #use the TWh value from each mix 
 
 
 vs <- getValues(solar_raster) #get the values from your energy potential map
 
 ## subset those values according to your sample -- get a randomized list of values from your energy potential raster with greater potential more likely to be chosen first
 vs <- vs[samp_s]
 
 vs.sum <- cumsum(vs) #get a running total of summed energy delivered
 stop <- min(which(vs.sum>thresh_s)) #stop once threshold is met
  
 selectrast_s<-solar_raster #make a new blank raster for final selected cells
 selectrast_s[]<-0 
 
 selectrast_s[samp_s[1:stop]]<-1 #set all cells in your sample (up to where you achieve energy target to 1)
  
## see how it all looks!
# plot(selectrast_s)

 #mask the output of the solar sample raster from the probability raster for hydro
 #set the masked values to zero
 layer1_s_h <- mask(probrast_h, selectrast_s, maskvalue = 1, updatevalue=0) 
 
 ###### Now hydro (h) - since this is the next raster we have that works
 
  xh<-getValues(probrast_h) #get all the values in your probability raster as a vector
 
 xh[is.na(xh)]<-0 #set NA cells in raster to zero
 
  ncellh <- length(which(xh!=0))
 
 ### sample from a list of cell IDs from 1:total cells in raster -- size is number in sample, and prob is the probability of selecting any cell (values from probrast)
 samp_h<-sample(nrow(probrast_h)*ncol(probrast_h), size=ncellh, prob=xh)
 
 samprast_h<-probrast_h #make a new raster to see what was sampled
 
 samprast_h[samp_h]<-1 #brackets say turn all the cell IDs identified by 'samp' into 1

 thresh_h <- thresh.hydro #use the TWh value from each mix 
 
 vh <- getValues(hydro_raster) #get the values from your energy potential map
 
 ## subset those values according to your sample -- get a randomized list of values from your energy potential raster with greater potential more likely to be chosen first
 vh <- vh[samp_h]
 
 vh.sum <- cumsum(vh) #get a running total of summed energy delivered
 #stop <- min(which(vh.sum>thresh_h)) #stop once threshold is met
 stop <-10000 
 
 selectrast_h<-hydro_raster #make a new blank raster for final selected cells
 selectrast_h[]<-0 
 
 selectrast_h[samp_h[1:stop]]<-2 #set all cells in your sample (up to where you achieve energy target to 1)
 
 #Hydro = 2, wind = 3, 4 = biofuels
  
## see how it all looks!
 #plot(selectrast_h)

 #mask the output of the hydro sample raster from the probability raster for wind
 #set the masked values to zero
# layer2_h_s <- mask(probrast_w, layer1_s_h, maskvalue = 0) 
 
 
 combined <- selectrast_h+selectrast_s
 plot(combined)
 
 writeRaster(combined,paste0("C:/Users/Jamie K/Documents/UCSB/Fall 2019/ESM 240 Climate Change Biology/Term Paper/team_paper/final/final_map",i,".tif"),overwrite=TRUE)
 
 
}


```









